### 2018-10-01 Project2/report/Project2.tex
Starting work on quantum dots, two electron

\documentclass{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{graphicx}
\usepackage{float}
\usepackage{amsmath}
\usepackage{epsfig,floatflt}
\setcitestyle{square}

\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\SetMathAlphabet{\mathcal}{bold}{OMS}{cmsy}{b}{n}

\begin{document}

\title{Project 2}

\author{Stig-Nicolai Foyn, Christer Dreierstad}

\email{stignicf@student.matnat.uio.no, chrisdre@student.matnat.uio.no}

\altaffiltext{1}{Institute of Physics, University of
  Oslo, P.O.\ Box 1029 Blindern, N-0315 Oslo, Norway}

\begin{abstract}
\end{abstract}
\keywords{buckling beam --- quantum dots --- harmonic oscillator- and Coulomb potentials --- methods: scaling of differential equations, Jacobi's method}

\section{Introduction}
\label{sec:introduction}

\section{Method}
\label{sec:method}
The method presented will show how to solve an eigenvalue problem using Jacobi's method. The first problem is a classical one involving a beam fastened in both ends. Further the method is transformed to a quantum mechanical three dimensional quantum dot problem, starting with one electron and moving over to two electrons.

\subsection{Jacobi's method for solving an eigenvalue problem}
Solving the eigenvalue problem with Jacobi's method means solving an equation on the form
%
\begin{gather}\label{eq:B=SAS}
    \bold{S}^T\bold{A}\bold{S} = \bold{B},
\end{gather}
%
where $\bold{S}$ is an orthogonal matrix, so it has the property $\bold{S}^{-1} = \bold{S}^T$. The resulting matrix $\bold{B}$ will consist of the eigenvalues along the diagonal. The matrix $\bold{S}$ is such that the elements along the diagonal are 1, and zero elsewhere, except four elements which are such that their position are decided by the largest element of the off-diagonal of $\bold{A}$. The largest element of $\bold{A}$ is found at a position $(k,l)$, and these indices are used to define the position of the mentioned four elements of $\bold{S}$, which are such that $s_{kk} = s_{ll} = \cos\theta$ and $s_{kl} = -s_{lk} = -\sin\theta$. This is considered as a rotation of an angle $\theta$ in the n-dimensional Euclidean space. When solving \eqref{eq:B=SAS} the resulting equations to solve are
%
\begin{align*}
    &b_{ii} = a_{ii}, i \neq k, i \neq l \\
    &b_{ik} = a_{ik}c - a_{il}s, i\neq k, i\neq l \\
    &b_{il} = a_{il}c + a_{ik}s, i \neq k, i\neq l \\
    &b_{kk} = a_{kk} c^2 - 2a_{kl}cs+ a_{ll}s^2 \\
    &b_{ll} = a_{ll} c^2 + 2a_{kl}cs a_{kk}s^2 \\
    &b_{kl} = \left(a_{kk} - a_{ll}\right)cs + a_{kl}\left(c^2 - s^2\right),
\end{align*}
where $c = \cos\theta$ and $s = \sin\theta$. The main goal is to reduce the off-diagonal elements of $\bold{A}$ to zero in such a way that $\bold{B}$ has the eigenvalues along the diagonal, the reduction of the elements is done by choosing the angle $\theta$. We make a requirement that the off-diagonal elements of $\bold{B}$ is zero;
%
\begin{align*}
    b_{kl} &= 0 = \left(a_{kk} - a_{ll}\right)cs + a_{kl} \left(c^2 - s^2\right),
\end{align*}
%
which can then be written as
%
\begin{gather*}
    \frac{\left(a_{ll} - a_{kk}\right)}{a_{kl}}cs = \left(c^2 - s^2\right).
\end{gather*}
%
Defining $\tau = \left(a_{ll} - a_{kk}\right)/2a_{kl}$ and $\tan\theta = t = s/c$, and multiplying the equation above by $1/c^2$ we obtain a quadratic equation 
%
\begin{gather*}
    t^2 + 2\tau t - 1 = 0,
\end{gather*}
%
which is used to define the best fit $\theta$ to reduce the off-diagonal elements of $\bold{A}$. The solution to the quadratic equation is $t = -\tau \pm \sqrt{1 + \tau^2}$. If $\tau$ becomes large, the numerical accuracy will drop and $t \approx -\tau \pm \tau$. By the following transformation
%
\begin{align*}
    t_{\pm} &= -\tau \pm \sqrt{1+\tau^2}\left(\frac{\tau \pm \sqrt{1+\tau^2}}{\tau \pm \sqrt{1+\tau^2}}\right) \\
    &= \frac{-\tau^2 + \left(1 + \tau^2\right)}{\tau \pm \sqrt{1 + \tau^2}} = \frac{1}{\tau \pm \sqrt{1+\tau^2}}.
\end{align*}
%
and by splitting the equation such that when $\tau < 0$ we choose the negative sign in front of the square root the aforementioned issue is avoided. The final equations for t are
%
\begin{align*}
    t_+ = \frac{1}{\tau + \sqrt{1 + \tau^2}}, \tau \geq 0 \\
    t_- = \frac{1}{\tau - \sqrt{1 + \tau^2}}, \tau < 0
\end{align*}
%
From the definition of $t = \tan\theta$, s can be expressed as
%
\begin{align*}
    s = \sin\theta = \tan\theta \cos\theta = tc,
\end{align*}
and c can be expressed as
\begin{align*}
    c &= \cos\theta = \frac{\cos\theta}{\sqrt{\cos^2\theta + \sin^2\theta}} = \frac{1}{\sqrt{1 + \tan^2\theta}}  = \frac{1}{\sqrt{1 + t^2}}.
\end{align*}

\subsubsection{Numerical implementation}
When implementing Jacobi's method numerically, $\bold{B}$ is not explicitly created, instead the elements of $\bold{A}$ are changed during each iteration. This allows the program to run with large matrices (large n), without taking up more memory thasn needed. This means the left hand side of the six previous equations are changed so that they are valid for the elements of $\bold{A}$ instead of $\bold{B}$, meaning $b_{ii} \rightarrow a_{ii}$, $b_{ik} \rightarrow a_{ik}$ and so forth.

///EIGENVECTORS////

\subsection{Buckling beam}
Considering a beam that is fastened in both ends which moves up and down when a force F is exerted on it can be described by a differential equation
%
\begin{gather}\label{eq:du2/dx2}
    R \frac{d^2 u(x)}{dx^2} = -Fu(x),
\end{gather}
%
where $u(x)$ is the displacement of the beam in the y-direction and $R$ is a constant that is defined by the properties of the beam, such as the rigidity. Since the beam, with length L, is fastened in both ends, the displacement at the end points $u(0) = u(L) = 0$, so we have Dirichlet boundary condition as in the previous paper \cite{1}, and the sum of outer forces exerted on the beam is zero. Introducing the dimensionless variable $\rho = x/L \Rightarrow x = \rho L$, where $\rho \in [0,1]$. Introducing this into Eq. \eqref{eq:du2/dx2} gives
%
\begin{gather*}
    \frac{R}{L}\frac{d^2 u(\rho)}{d\rho^2} = - FLu(\rho)
    \Rightarrow \frac{d^2 u(\rho)}{d\rho^2} = -\frac{FL^2}{R}u(\rho).
\end{gather*}
%
Substituting $\lambda = FL^2/R$ transforms the problem to an eigenvalue problem, with discretized $\lambda$'s as eigenvalues with corresponding eigenvectors. The methods presented in the previous paper regarding the solution of a differential equation will be applied to this problem. The left hand size of the equation to solve,
%
\begin{gather}\label{eq:du/dx=lambda}
    \frac{d^2 u(\rho)}{dx^2} = -\lambda u(\rho)
\end{gather}
%
can be approximated as
%
\begin{gather*}
    \frac{d^2 u(\rho)}{dx^2} = \frac{u(\rho + h) - 2u(\rho) + u(\rho - h)}{h^2} +  \mathcal{O}(h^2),
\end{gather*}
%
where the step size $h = \left(\rho_{max} - \rho_{min}\right)/n$, which is the total length of the beam divided by the resolution of the calculations n. The maximum and minimum value of the dimensionless length variable $\rho$ is at the end points of the beam, such that $\rho_{max} = 1$ and $\rho_{min} = 0$. Following the method from the previous paper the discretized equation to solve is
%
\begin{gather*}
    -\frac{u_{i+1} - 2u_i + u_{i-1}}{h^2} = \lambda u_i.
\end{gather*}
Which by the method of the previous paper can be transferred to a matrix eigenvalue problem;
%
\begin{equation*}
    \begin{bmatrix} d& a & 0   & 0    & \dots  &0     & 0 \\
                                a & d & a & 0    & \dots  &0     &0 \\
                                0   & a & d & a  &0       &\dots & 0\\
                                \dots  & \dots & \dots & \dots  &\dots      &\dots & \dots\\
                                0   & \dots & \dots & \dots  &a  &d & a\\
                                0   & \dots & \dots & \dots  &\dots       &a & d\end{bmatrix} 
                                 \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} = \lambda \begin{bmatrix} u_1 \\ u_2 \\ u_3 \\ \dots \\ u_{N-2} \\ u_{N-1}\end{bmatrix} . 
\label{eq:matrixse} 
\end{equation*}
%
Where a and b are defined as $a = -1/h^2$ and $b = 2/h^2$. For this specific eigenvalue problem the analytical eigenvalues are
%
\begin{gather}\label{eq:eigenvals_analytical}
    \lambda_j = d + 2a\cos\left(\frac{j\pi}{n+1}\right)
\end{gather}
for $j = 1,2,...,n-1$.

\subsection{Quantum dots, one electron}
The method derived in the section for the buckling beam is general for a problem that can be described by a differential equation on the form of Eq. \eqref{eq:du/dx=lambda}. The method can be extended to a quantum mechanical problem, if the problem can be transformed into a differential equation as mentioned. The radial part of the Schr$\ddot{\mathrm{o}}$dinger equation with the orbital angular momentum number $l = 0$ is
%
\begin{equation*}
    -\frac{\hbar^2}{2m}\left(\frac{1}{r^2}\frac{d}{dr}r^2\frac{d}{dr}\right)R(r) + V(r)R(r) = ER(r),
\end{equation*}
%
where $r \in [0,\infty)$, the potential $V(r)$ is the harmonic oscillator potential $V(r) = (1/2)kr^2$, where $k = mw^2$ and $\omega$ is the oscillator frequency. The allowed energies are 
%
\begin{equation*}
    E_{n0} = \left(2n + \frac{3}{2}\right),
\end{equation*}
%
for  n = 0, 1, 2,.... Substituting $R(r) = u(r)/r$:
%
\begin{equation*}
    -\frac{\hbar^2}{2m}\frac{1}{r^2}\frac{d}{dr}r^2 \frac{d}{dr}\frac{u(r)}{r} + V(r)\frac{u(r)}{r} = E\frac{u(r)}{r}.
\end{equation*}
%
Evaluating the derivatives in the above equation:
%
\begin{align*}
    \frac{1}{r^2}\frac{d}{dr}r^2\frac{d}{dr}\frac{u(r)}{r} &=  \frac{1}{r^2}\frac{d}{dr}r^2\left(\frac{\frac{du}{dr}r - u}{r^2}\right) \\
    &= \frac{1}{r^2}\frac{d}{dr}\left(\frac{du}{dr}r - u\right) \\
    &= \frac{1}{r^2}\left(\frac{d^2u}{dr^2}r + \frac{du}{dr} - \frac{du}{dr}\right) \\
    &= \frac{1}{r}\frac{d^2u}{dr^2},
\end{align*}
%
Inserting the above and multiplying by r yields
%
\begin{equation*}
    -\frac{\hbar^2}{2m}\frac{d^2u}{dr^2} + V(r)u(r) = Eu(r).
\end{equation*}
%
where the boundary conditions of $u$ are $u(0) = u(\infty) = 0$. Defining a dimensionless variable $\rho = r/\alpha$ transforms the potential to $V(\rho) = (1/2)k\alpha^2\rho^2$, and the Schr$\ddot{\mathrm{o}}$dinger's equation becomes
%
\begin{equation*}
    -\frac{\hbar}{2m\alpha^2}\frac{d^2u(\rho)}{d\rho^2} + \frac{1}{2}k\alpha^2\rho^2u(\rho) = Eu(\rho).
\end{equation*}
%
Multiplying the above equation by $2m\alpha^2 / \hbar^2$
%
\begin{equation*}
    -\frac{d^2u(\rho)}{d\rho^2} + \frac{mk}{\hbar^2}\alpha^4 \rho^2u(\rho) = \frac{2m\alpha^2}{\hbar^2}Eu(\rho).
\end{equation*}
%
Where the factor in front of $u(\rho$ on the right hand side is defined as
%
\begin{equation*}
    \lambda = \frac{2m\alpha^2}{\hbar^2}E,
\end{equation*}
%
which corresponds to the eigenvalues of the problem. To transform the equation to a form such that it can be solved as an eigenvalue problem, $\alpha$ is fixed such that
%
\begin{equation*}
    \frac{mk}{\hbar^2}\alpha^4 = 1
    \Rightarrow \alpha = \left(\frac{\hbar^2}{mk}\right)^{1/4}.
\end{equation*}
%
The rewritten expression for the Schr$\ddot{\mathrm{o}}$dinger equation is then
%
\begin{equation*}
    -\frac{d^2}{d\rho^2}u(\rho) + \rho^2u(\rho) = \lambda u(\rho),
\end{equation*}
%
which has known analytical eigenvalues for the ground state (with $l=0)$ $\lambda_{n=0} = 3$ and for the two first excited states $\lambda_1 = 7, \lambda_2 = 11, \lambda_3 = 15$.

By the same method as for the buckling beam we can discretize the Schr$\ddot{\mathrm{o}}$dinger equation:
%
\begin{equation*}
    -\frac{u_{i+1} - 2u_i + u_{i-1}}{h^2} + \rho_i^2u_i = \lambda u_i,
\end{equation*}
%
where the second term on the left comes from the $\rho_i^2 = V_i$, the harmonic oscillator potential. The step length is as previous $h = \left(\rho_n - \rho_0\right)/n$, where $\rho_n = \rho_{max} \rightarrow \infty$. In the calculations $\rho_n$ is approximated by a value that is large enough such that the wave function is zero at that point, and the physics are of no further interest. Evaluating the equation as a tridiagonal matrix the diagonal elements are defined as
%
\begin{equation*}
    d_i = \frac{2}{h^2} + V_i,
\end{equation*}
%
and the upper and lower diagonal elements as
%
\begin{equation*}
    a_i = -\frac{1}{h^2},
\end{equation*}
%
and the final discretized Schr$\ddot{\mathrm{o}}$dinger equation is
%
\begin{equation*}
    d_iu_i + e_{i-1}u_{i-1} + e_{i+1}u_{i+1} = \lambda u_i,
\end{equation*}
%
which is on the same form as the equation for the buckling beam, and the problem can be solved by the established Jacobi's method.

\subsection{Quantum dots, two electrons}

\section{Results}
\label{sec:results}

\subsection{Jacobi's method}

Comparing the results from Jacobi's method and the analytic results from the provided in [SOURCE:Poject 2], we achieve sufficient numerical precision even for variyng sizes between $n = 5$ and $n = 400$.
%
\begin{deluxetable}{lcc}
%\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Table comparing CPU-time between Armadillo's built in function (eigsym) and Jacobi's method, as visible from this data Jacobi's method increases in time spent considerably faster than that of Armadillo's function.}
\tablecolumns{4}
\tablehead{Size n  & Armadillo & Jacobi's method}
\startdata
10 & 0.000262s & 0.000155s \\
50 & 0.000915s & 0.007585s \\
100 & 0.005636s & 0.106488s \\
200 & 0.018883s & 1.53787s \\
300  & 0.04241s & 8.85018s \\
400  & 0.085901s & 30.4565s
\enddata
\end{deluxetable}
%
%
\begin{figure}[t]
\mbox{\epsfig{figure=iterationspersize15.pdf,width=\linewidth,clip=}}
\caption{Logarithmic plot of the number of iteration needed as function of the matrix dimentionality n, to illustrate the quadratic dependency we have plotted $y = 2log_{10}(n)$, which in the plot is seen as the dotted $y=2x$ line.}
\label{fig:fig1}
\end{figure}
%
\subsection{Quantum dots, single electron}
Analyzing how the number of integration points $n$ and the endpoint $\rho_{n}$, and how many integration points is needed to achieve four digit accuracy in the eigenvalues as a result from using Jacobi's method on a Toepliz matrix with quantum harmonic oscillator potential term. For $\rho_{n} = 5$, the size n required to achieve four leading digits is $n = 380$, which returns the value $E_{00} = 2.99995$ .

%
CONSIDERING ADDING PLOTS OF INTEGRATION POINTS AND ENDPOINT ANALYSIS.
%

\subsection{Quantum dots, two electrons}

\begin{figure}[t]
\mbox{\epsfig{figure=GSEperOMGr15.pdf,width=\linewidth,clip=}}
\caption{Plot of the ground state energy as function of $\omega_{r}$, it seemingly increases linearly with slope $a = 3.4369$, with error $da = 0.0511$}
\label{fig:fig2}
\end{figure}
%
\begin{deluxetable}{lcc}
%\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Table comparing ground stat analytic energy eigenvalues found in [SOURCE:] and numerical energy eigenvalues, the analytic eigenvalues displayed in the table of the source are halved, $\epsilon ' = 0.6250$, for $\omega_{r} = 0.25$ and $\epsilon ' = 0.1750$ for $\omega_{r} = 0.05$.}
\tablecolumns{4}
\tablehead{$\omega_{r}$ & Analytic $2\epsilon '$ & Numerical $E_{00}$}
\startdata
0.25 & 1.250 & 1.24999 \\
0.05 & 0.350 & 0.349999
\enddata
\end{deluxetable}
%
\begin{figure}[t]
\mbox{\epsfig{figure=ProbfuncperRhons15.pdf,width=\linewidth,clip=}}
\caption{Plot of probability distribution of the two electron quantum dot in a harmonic oscillator potential for different $\omega_{r}$, as well as for no Coulomb interaction. The probability is plotted as a function of the dimensionless relative particle position $\rho = \rho_n - \rho_0$. The maximum value of the $\rho_n$ is an approximation of an infinite value, and is selected as the value where the wave function has vanished}
\label{fig:fig3}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}

%We have managed to reduce a 3 dimentional two body problem to a analytically solvable problem by changing coordinate system from euclidian to radial, and analyizing CM-system(?)

%\begin{figure}[t]
%
%\mbox{\epsfig{figure=filename.eps,width=\linewidth,clip=}}
%
%\caption{Description of figure -- explain all elements, but do not
%draw conclusions here.}
%\label{fig:figure_label}
%\end{figure}

\begin{deluxetable}{lccc}
%\tablewidth{0pt}
\tablecaption{\label{tab:results}}
\tablecomments{Summary of main results.}
\tablecolumns{4}
\tablehead{Column 1  & Column 2 & Column 3 & Column 4}
\startdata
Item 1 & Item 2 & Item 3 & Item 4
\enddata
\end{deluxetable}

\begin{acknowledgements}
\end{acknowledgements}

\begin{thebibliography}{9}

\bibitem{1}
Stig-Nicolai Foyn, Christer Dreierstad
\textit{Project 1}

\end{thebibliography}

\end{document}